---
title: "EMR RK Analysis Script"
output: html_document
---

```{r, warning=F, message=F}
library(rebus)
library(cowplot)
library(gridExtra)
library(readxl)
library(RColorBrewer)
library(ggrepel)
library(tidyverse)

theme_set(theme_bw() + theme(axis.text = element_text(colour = "black", size = 14.5),
                             axis.title = element_text(colour = "black", face = "bold", size = 15.5),
                             strip.background = element_blank(),
                             panel.border = element_rect(size = .5, color = "black"),
                             strip.text = element_text(face = "bold", size = 12)))
```

# Sorbent optimization
```{r, warning=F, message=F}
path = "/Users/Boyuan/Desktop/My publication/17. EMR Adipose RK metabolites/data RK lipid EMR validation 3th ,Zhiya 081920.xlsx"

color.cmpd = colorRampPalette(brewer.pal(8, "Dark2"))(26) %>% sort() %>% sort()

d.opt.sorbent = read_excel(path, sheet = "sorbent opt")

d.opt.sorbent = d.opt.sorbent %>%
  gather(-c(`Data File`, sorbent, `Acq. Date-Time`), key = compound, value = resp)

x = d.opt.sorbent %>% filter(sorbent == "ctrl") %>%
  group_by(compound) %>%
  summarise(resp.ctrl.mean = mean(resp),
            resp.ctrl.sd = sd(resp))

y = d.opt.sorbent %>% filter(sorbent != "ctrl") %>%
  group_by(sorbent, compound) %>%
  summarise(resp.mean = mean(resp), resp.sd = sd(resp))

d.opt.sorbent.recovery = y %>% left_join(x, by = c("compound")) %>%
  mutate(recovery.mean = resp.mean / resp.ctrl.mean * 100,
         recovery.sd = sqrt((resp.sd/resp.mean)^2 + (resp.ctrl.sd/resp.ctrl.mean)^2) * recovery.mean  )
d.opt.sorbent.recovery

d.opt.sorbent.recovery$sorbent = d.opt.sorbent.recovery$sorbent %>% 
  factor(levels = c("Sorbent free", "100 NP (big size)", "100 NP", "75 NP-25 RP", "50 NP-50 RP", "25 NP-75 RP", "100 RP"), ordered = T)

# low recovery compounds without sorbent
l = d.opt.sorbent.recovery %>% filter(sorbent == "Sorbent free") %>%
  filter(recovery.mean < 30)


cmpd.lowRecovery = c("RK-Me", "PhLiAce", "3,4-DHPPA", "3,4-DHPAA", "CA", "4-HBOH")

dg = .3
plt.sorbentOptimization = d.opt.sorbent.recovery %>%
  filter(!compound %in% cmpd.lowRecovery ) %>%
  ggplot(aes(x = sorbent, y = recovery.mean, color = compound, fill = compound)) +
  
  geom_boxplot(data = d.opt.sorbent.recovery, outlier.alpha = 0, 
               width = .8, fill = "snow1",
               aes(group = sorbent)) +
  
  geom_errorbar(aes(ymin = recovery.mean - recovery.sd,
                    ymax = recovery.mean + recovery.sd),
                position = position_dodge(dg), size = .1, width = 1) +
  geom_point(position = position_dodge(dg), shape = 21, fill = "white", size = 2) +
  
  theme(legend.position = "none", panel.grid = element_blank(),
        axis.text = element_text(size = 11, colour = "black"),
        axis.title = element_text(size = 12, colour = "black")) +
  scale_y_continuous(breaks = seq(0, 150, by = 10)) +
  coord_cartesian(ylim = c(0, 110)) +
  
  # threshold
  geom_segment(aes(x = .5, xend =  7.5, y = 30, yend = 30), 
               linetype = "dashed", size = 1, color = "firebrick") +
  
  # add special note to low-recovery compound
  
  geom_line(data = d.opt.sorbent.recovery %>% filter(compound %in% cmpd.lowRecovery),
            aes(group = compound), position = position_dodge(dg), linetype = "dotted") +
  
  geom_errorbar(data = d.opt.sorbent.recovery %>% filter(compound %in% cmpd.lowRecovery),
                aes(ymin = recovery.mean - recovery.sd,
                    ymax = recovery.mean + recovery.sd),
                position = position_dodge(dg), size = .1, width = .1) +
  
  geom_point(data = d.opt.sorbent.recovery %>% filter(compound %in% cmpd.lowRecovery),
             aes(group = compound), position = position_dodge(dg),
             shape = 21, fill = "white", size = 2, stroke = 1) +
  
  # fill with transparent color
  geom_point(data = d.opt.sorbent.recovery %>% filter(compound %in% cmpd.lowRecovery),
             aes(group = compound), position = position_dodge(dg),
             shape = 21, size = 2, alpha = .6) +
  
  geom_text_repel(data = d.opt.sorbent.recovery %>% filter(compound %in% cmpd.lowRecovery),
                  aes(label = compound), size = 3, fontface = "bold",
                  position = position_dodge(dg)) +
  labs(y = "Recovery (%)", x = "Sorbent") +
  scale_color_manual(values = color.cmpd) +
  scale_fill_manual(values = color.cmpd) 

# 7.5 X 8.9 Mac screen
```


```{r, warning=F, message=F, fig.width=10}
plt.sorbentOptimization
```


```{r, warning=F, message=F}
d.opt.sorbent.recovery %>%
  filter(compound %in% c("RK-Me", "PhLiAce", "4-HBOH")) %>%
  select(sorbent, compound, recovery.mean) %>% spread(sorbent, recovery.mean)

d.opt.sorbent.recovery$compound %>% n_distinct()



# Effect of sorbent mass
d.opt.sorbent.mass = read_excel(path, sheet = "sorbent mass")
d.opt.sorbent.mass = d.opt.sorbent.mass %>% 
  gather(-c(`Data File`, code, `Acq. Date-Time`, `sorbent mass`), key = compound, value = resp) %>%
  group_by(compound) %>%
  mutate(resp.relative = resp / mean(resp)) 

d.opt.sorbent.mass = d.opt.sorbent.mass %>%
  group_by(`sorbent mass`, compound) %>%
  summarise(resp.relative.mean = mean(resp.relative), resp.relative.sd = sd(resp.relative)) 
d.opt.sorbent.mass

plt.sorbent.mass = d.opt.sorbent.mass %>%
  ggplot(aes(x = `sorbent mass`, y = resp.relative.mean, color = compound)) +
  geom_boxplot(aes(group = `sorbent mass`), outlier.alpha = 0) +
  geom_errorbar(aes(ymin = resp.relative.mean - resp.relative.sd,
                    ymax = resp.relative.mean + resp.relative.sd),
                position = position_dodge(.5), size = .2, width = 1.2) +
  geom_point(position = position_dodge(.5), shape = 21, fill = "white") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(.5, 1.6)) +
  scale_y_continuous(breaks = seq(.6, 1.6, .2)) +
  labs(y = "Relative peak intensity", caption = "Reconstituted in 100 uL 60% MeOH 0.1% FA before LCMS analysis")

plt.sorbent.mass


# <>0^%@#$!)(*&^$##&$)^$(^#!~#$)# <>0^%@#$!)(*&^$##&$)^$(^#!~#$)# <>0^%@#$!)(*&^$##&$)^$(^#!~#$)# <>0^%@#$!)(*&^$##&$)^$(^#!~#$)
```


# Accuracy validation
```{r, warning=F, message=F}
# Validation
d.conc = read_excel(path, sheet = "calculated concentration")
d.conc = d.conc %>% 
  gather(-c(`Data File`, Name, `Acq. Date-Time`), key = compound, value = conc) %>%
  select(-`Acq. Date-Time`)

# blank concentration
d.conc.blk = d.conc %>% filter(Name == "Adi + Enz") %>% 
  select(-c(Name, `Data File`)) %>%
  group_by(compound) %>%
  summarise(conc.blk.mean = mean(conc, na.rm = T),
            conc.blk.sd = sd(conc, na.rm = T))

# total measured conc. (spiked + background)
d.ABCD = d.conc %>% filter(Name %in% c("A", "B", "C", "D")) %>%
  group_by(compound, Name) %>%
  summarise(conc.all.mean = mean(conc, na.rm = T),
            conc.all.sd = sd(conc, na.rm = T))

# subract background
d.ABCD = d.ABCD %>% left_join(d.conc.blk, by = "compound") %>%
  mutate(conc.spk.mean = conc.all.mean - conc.blk.mean,
         conc.spk.sd = (conc.all.sd^2 + conc.blk.sd^2) %>% sqrt()) %>%
  select(compound, Name, contains("spk"))


# Expected spiked amount
d.expect = read_excel(path, sheet = "stock solution", range = "B2:O28") %>%
  select(compound, A, B, C, D)

d.expect = d.expect %>%
  gather(-compound, key = Name, value = conc.expected)


# Compute Accuracy
d.ABCD = d.ABCD %>% left_join(d.expect, by = c("compound", "Name")) %>%
  mutate(Accuracy.mean = conc.spk.mean / conc.expected * 100,
         Accuracy.sd = conc.spk.sd / conc.expected * 100) %>% 
  # add spike concentration level
  ungroup() %>%
  mutate(level = rep(c("2000", "500", "100", "20"), time = d.ABCD$compound %>% n_distinct()),
         level = factor(level, levels = c("2000", "500", "100", "20"), ordered = T))  

# 3,4-DHPPA D level below limit of quantification
d.ABCD = d.ABCD %>% filter(! ((compound == "3,4-DHPPA") & (Name == "D")))


# accuracy box plot
dg = .4
```


```{r, warning=F, message=F}
plt.accuracy = d.ABCD %>%
  ggplot(aes(x = level, y = Accuracy.mean, color = compound, fill = compound)) +
  
  geom_boxplot(outlier.alpha = 0, aes(group = Name), fill = "grey", width = .8) +
  
  geom_errorbar(aes(ymin = Accuracy.mean - Accuracy.sd,
                    ymax = Accuracy.mean + Accuracy.sd ),
                size = .2, width = 1, position = position_dodge(dg)) +
  
  geom_point(position = position_dodge(dg), shape = 21, fill = "white", size = 4.5) +
  geom_point(position = position_dodge(dg), shape = 21, alpha = 0.2, size = 4.5) +
  
  theme(legend.position = "None", panel.grid = element_blank()) +
  scale_y_continuous(breaks = seq(0, 180, by = 20)) +
  
  geom_text_repel(data = d.ABCD %>% 
                    filter(Accuracy.mean > 140 | Accuracy.mean < 60 ),
                  aes(label = compound), size = 5, position = position_dodge(dg)) +
  
  coord_cartesian(ylim = c(0, 180)) +
  
  annotate("segment", x = .5, xend = 4.5, y = 120, yend = 120, 
           linetype = "dashed", size = .6, color = "firebrick") +
  annotate("segment", x = .5, xend = 4.5, y = 80, yend = 80, 
           linetype = "dashed", size = .6, color = "firebrick") +
  
  annotate("segment", x = .5, xend = 4.5, y = 100, yend = 100, 
           linetype = "dashed", size = .6, color = "darkgreen") +
  
  scale_color_manual(values = color.cmpd) +
  scale_fill_manual(values = color.cmpd) +
  labs(y = "Accuracy", x = "Spiked concentration of final injection (ng/mL)")
```


```{r, warning=F, message=F, fig.width=8}
plt.accuracy
```


```{r, warning=F, message=F}
# annotation standard: Accuracy.mean > 140 | Accuracy.mean < 60 | Accuracy.sd > 50

d.ABCD %>% filter(Accuracy.sd > 50) 


# <>)*$#*(#@!$&<>*&%$#@!@#^*^# <>)*$#*(#@!$&<>*&%$#@!@#^*^# <>)*$#*(#@!$&<>*&%$#@!@#^*^# <>)*$#*(#@!$&<>*&%$#@!@#^*^
```


# Stepwise recovery validation
```{r, warning=F, message=F}
# Recovery at each step
d.area = read_excel(path, sheet = "peak area")
d.area = d.area %>% 
  gather(-c(`Data File`, Name, `Acq. Date-Time`), key = compound, value = area)

d.area.copy = d.area # make a copy for later use with comparison with internal standard

# background 
d.area.blk = d.area %>% filter(Name == "Adi + Enz") %>% 
  group_by(compound) %>%
  summarise(area.blk.mean = mean(area, na.rm = T),
            area.blk.sd = sd(area, na.rm = T))

# total area at each step ( spiked + background )
d.area.steps = d.area %>% 
  filter(Name %in% c("B", "C", "PX", "PC", "PR", "PR neat", "no silica gel", "PR neat")) %>%
  group_by(compound, Name) %>%
  summarise(area.mean = mean(area, na.rm = T),
            area.sd = sd(area, na.rm = T))
#good

# subtract background
d.area.steps = d.area.steps %>% left_join(d.area.blk, by = "compound") %>%
  mutate(area.spk.mean = area.mean - area.blk.mean,
         area.spk.sd = (area.sd^2 + area.blk.sd^2) %>% sqrt()) %>%
  select(compound, Name, contains("spk"))
d.area.steps  

# spread apart steps into separate columns
x = d.area.steps %>% select(-area.spk.sd) %>% 
  spread(key = Name, value = area.spk.mean)
colnames(x) = c("compound", colnames(x)[-1] %>% str_c(".mean"))

y = d.area.steps %>% select(-area.spk.mean) %>% 
  spread(key = Name, value = area.spk.sd)
colnames(y) = c("compound", colnames(y)[-1] %>% str_c(".sd"))
y

d.area.steps = left_join(x, y, by = "compound")


# recovery at each steps (now background subtracted)
d.recovery = d.area.steps %>%
  mutate(
    # extraction (relative to spike amount of post extraction or PX)
    rec.extraction.mean = B.mean / PX.mean  * 100, 
    rec.extraction.sd = sqrt((PX.sd / PX.mean)^2 + (B.sd / B.mean)^2) * rec.extraction.mean,
    
    # EMR clean up (relative to spike amount of post clean up or PC)
    rec.EMR.mean = PX.mean / PC.mean * 100,
    rec.EMR.sd = sqrt((PX.sd / PX.mean)^2 + (PC.sd / PC.mean)^2)* rec.EMR.mean,
    
    # SpeedVac drying & reconstitution (relative to spike amount of post reconstitution or PR)
    rec.DryReconstitute.mean = PC.mean / PR.mean * 100, 
    rec.DryReconstitute.sd = sqrt((PC.sd / PC.mean)^2 + (PR.sd / PR.mean)^2) * rec.DryReconstitute.mean,
    
    # # add silica gel effect (relative to without, using spike level C)
    # sorbent.mean = C.mean  / `no silica gel.mean`, # (improvenment fold), with improvement > 1
    # sorbent.sd = sqrt((`no silica gel.sd` / `no silica gel.mean`)^2 + (C.sd / C.mean)^2) * sorbent.mean,
    
    # Overall recovery 
    rec.Overall.mean = B.mean / PR.mean * 100,
    rec.Overall.sd = sqrt((B.sd / B.mean)^2 + (PR.sd/ PR.mean)^2) * rec.Overall.mean,
    
    # matrix effect to ESI
    matrixESI.mean = PR.mean / `PR neat.mean` * 100,
    matrixESI.sd = sqrt((PR.sd / PR.mean)^2 + (`PR neat.sd` / `PR neat.mean`)^2) * matrixESI.mean,
    
    # Overall processing efficiency
    process.eff.mean = B.mean / `PR neat.mean` * 100,
    process.eff.sd = sqrt((B.sd / B.mean)^2 + (`PR neat.sd` / `PR neat.mean`)^2) * process.eff.mean) %>%
  
  select(compound, contains("rec"), contains("ESI"), contains("process"))


x = d.recovery %>%
  select(compound, contains("mean")) %>%
  gather(-compound, key = step, value = mean) %>%
  mutate(step = str_remove(step, pattern = ".mean"))

y = d.recovery %>%
  select(compound, contains("sd")) %>%
  gather(-compound, key = step, value = sd) %>%
  mutate(step = str_remove(step, pattern = ".sd"))
y

d.recovery.tidy = x %>% left_join(y, by = c("compound", "step"))

# compound wise
# d.recovery.tidy %>%
#   ggplot(aes(x = compound, y = mean, color = step)) +
#   geom_point(position = position_dodge(dg)) +
#   coord_flip()

# overview
d.recovery.tidy$step = d.recovery.tidy$step %>% str_replace(pattern = "rec.extraction", replacement = "Extraction")
d.recovery.tidy$step = d.recovery.tidy$step %>% str_replace(pattern = "rec.EMR", replacement = "Enhanced matrix removal")
d.recovery.tidy$step = d.recovery.tidy$step %>% str_replace(pattern = "rec.DryReconstitute", replacement = "Dry & reconstitute")
d.recovery.tidy$step = d.recovery.tidy$step %>% str_replace(pattern = "rec.Overall", replacement = "Overall recovery")
d.recovery.tidy$step = d.recovery.tidy$step %>% str_replace(pattern = "matrixESI", replacement = "Matrix effect (ESI)")
d.recovery.tidy$step = d.recovery.tidy$step %>% str_replace(pattern = "process.eff", replacement = "Overall processing efficiency")

d.recovery.tidy$step = factor(d.recovery.tidy$step, levels = d.recovery.tidy$step %>% unique(), ordered = T)

plt.recovery.steps = d.recovery.tidy %>% 
  ggplot(aes(x = step, y = mean, color = compound, fill = compound)) +
  geom_boxplot(outlier.alpha = 0, fill = "grey", aes(group = step)) +
  
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),
                position = position_dodge(dg), 
                size = .2, width = 1) +
  geom_point(position = position_dodge(dg),
             shape = 21, size = 3, stroke = 1, fill = "white") +
  
  geom_point(position = position_dodge(dg),
             shape = 21, size = 3, stroke = 1, alpha = .2) +
  
  annotate("segment", x = .5, xend = 6.5, y = 100, yend = 100, 
           linetype = "dashed", size = .4, color = "darkgreen") +
  annotate("segment", x = .5, xend = 6.5, y = 80, yend = 80, 
           linetype = "dashed", size = .4, color = "firebrick") +
  
  coord_cartesian(ylim = c(20, 140)) +
  scale_y_continuous(breaks = seq(0, 200, by = 10)) +
  scale_color_manual(values = color.cmpd) +
  scale_fill_manual(values = color.cmpd) +
  theme(legend.position = "None") +
  labs(x = " ") +
  
  geom_text_repel(data = d.recovery.tidy %>% filter(mean > 110 | mean <= 50), 
                  aes(label = compound), position = position_dodge(dg), size = 3)
```


```{r, warning=F, message=F, fig.width=10}
plt.recovery.steps
```


```{r, warning=F, message=F}
# *<^%#^&*$#@!%$#^&*<^%#^&*$#@!%$#^&*<^%#^&*$#@!%$#^&*<^%#^&*$#@!%$#^&*<^%#^&*$#@!%$#^&*<^%#^&*$#@!%$#^&*<^%#^&*$#@!%$#^&
```


# Sorbent effect revisit
```{r, warning=F, message=F}
# sorbent improvement effect
# For this section the peak area should NOT subtract background since background is measured WITH sorbent. The with sorbent and without sorbent should be compared directly with background content included
d.area.sorbent = d.area %>% filter(Name %in% c("C", "no silica gel")) %>%
  group_by(compound, Name) %>%
  summarise(area.mean = mean(area),
            area.sd = sd(area))


x = d.area.sorbent %>% select(-area.sd) %>% spread(key = Name, value = area.mean)
colnames(x) = c("compound", "C.mean", "no silica gel.mean")

y = d.area.sorbent %>% select(-area.mean) %>% spread(key = Name, value = area.sd)
colnames(y) = c("compound", "C.sd", "no silica gel.sd")

d.area.sorbent = left_join(x, y, by = "compound") %>%
  mutate(sorbent.mean = C.mean / `no silica gel.mean`,
         sorbent.sd = C.sd / `no silica gel.sd`)

dg. =5
plt.sorbent.validate = d.area.sorbent %>% 
  ggplot(aes(x = 1, y = sorbent.mean, color = compound, fill = compound)) +
  geom_boxplot(outlier.alpha = 0, fill = "grey", aes(group = 1)) +
  
  geom_errorbar(aes(ymin = sorbent.mean - sorbent.sd, 
                    ymax = sorbent.mean + sorbent.sd),
                position = position_dodge(dg),
                size = .2, width = 1) +
  geom_point(position = position_dodge(dg),
             shape = 21, size = 3, stroke = 1, fill = "white") +
  
  geom_point(position = position_dodge(dg),
             shape = 21, size = 3, stroke = 1, alpha = .2) +
  
  annotate("segment", x = .5, xend = 1.5, y = 1, yend = 1, 
           linetype = "dashed", size = .4, color = "darkgreen") +
  scale_color_manual(values = color.cmpd) +
  scale_fill_manual(values = color.cmpd) +
  theme(legend.position = "None",
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())  +
  labs(x = "", y = "Fold increase of peak area with sorbent than without ") +
  
  geom_text_repel(data = d.area.sorbent %>% filter(sorbent.mean > 2),
                  aes(label = compound), position = position_dodge(dg), size = 3) +
  # scale_y_continuous(breaks = seq(0, 7, 1))+
  coord_cartesian(ylim = c(0, 6))
```


```{r, warning=F, message=F}
plt.sorbent.validate
```


```{r, warning=F, message=F}
#0*^$@$^!#%&*#0*^$@$^!#%&*#0*^$@$^!#%&*#0*^$@$^!#%&*#0*^$@$^!#%&*#0*^$@$^!#%&*#0*^$@
```

# ISTD-corrected stepwise recovery
```{r, warning=F, message=F}
# Recovery with internal standard correcting the loss effect

# IS area in each sample file
d.area.IS1 = read_excel(path, sheet = "IS1 peak area")
d.area.IS2 = read_excel(path, sheet = "IS2 peak area") 

d.area.IS = d.area.IS1 %>% 
  left_join(d.area.IS2 %>% select(-c(`Acq. Date-Time`, Name)),
            by = c("Data File"))

# d.area.IS[!complete.cases(d.area.IS), ]

# Compounds vs. IS
d.IS = read_excel(path, sheet = "IS")


# Compoute area corrected 
d.area.corrected = d.area.copy %>% 
  filter(Name %in% c("B", "PX", "PC", "PR", "PR neat")) %>%
  
  # subtract background on a sample wise manner
  left_join(d.area.blk, by = "compound") %>%
  mutate(spk.mean = area - area.blk.mean,
         spk.sd = area.blk.sd) %>%
  select(-c(area, contains("blk"))) %>%
  # which IS
  left_join(d.IS, by = "compound") 

# compounds using 4-HBA as IS
x = d.area.corrected %>% filter(IS == "4-HBA-d4") %>%
  left_join(d.area.IS %>% select(`Data File`, `4-HBA-d4`), by = "Data File") %>%
  rename(area.IS = `4-HBA-d4`)

# compounds using cinnamic acid as IS
y = d.area.corrected %>% filter(IS == "cinnamic-d6") %>%
  left_join(d.area.IS %>% select(`Data File`, `cinnamic-d6`), by = "Data File") %>%
  rename(area.IS = `cinnamic-d6`)

# combine 2 IS's
d.area.corrected = rbind(x, y)

d.area.corrected = d.area.corrected %>%
  select(-c(`Data File`, `Acq. Date-Time`)) %>%
  # now area corrected with area of IS, shown as the ratio
  mutate(ratio = spk.mean / area.IS,
         ratio.sd = spk.sd / area.IS) %>%
  select(-contains("spk"), -contains("IS"))

d.area.corrected = d.area.corrected %>% 
  group_by(Name, compound) %>%
  summarise(ratio.mean = mean(ratio, na.rm = T),
            ratio.sd = sd(ratio, na.rm = T))

# d.area.corrected[!complete.cases(d.area.corrected), ]


# spread columns apart
x = d.area.corrected %>% select(-ratio.sd) %>%
  spread(key = Name, value = ratio.mean)
colnames(x) = c("compound", colnames(x)[-1] %>% str_c(".mean") )

y = d.area.corrected %>% select(-ratio.mean) %>%
  spread(key = Name, value = ratio.sd)
colnames(y) = c("compound", colnames(y)[-1] %>% str_c(".sd") )

d.area.corrected = left_join(x, y, by = "compound")

# compute recovery corrected by IS
d.recovery.corrected = d.area.corrected %>%
  # the recovery code format is the same as above without correction
  mutate(
    # extraction (relative to spike amount of post extraction or PX)
    rec.extraction.mean = B.mean / PX.mean  * 100, 
    rec.extraction.sd = sqrt((PX.sd / PX.mean)^2 + (B.sd / B.mean)^2) * rec.extraction.mean,
    
    # EMR clean up (relative to spike amount of post clean up or PC)
    rec.EMR.mean = PX.mean / PC.mean * 100,
    rec.EMR.sd = sqrt((PX.sd / PX.mean)^2 + (PC.sd / PC.mean)^2)* rec.EMR.mean,
    
    # SpeedVac drying & reconstitution (relative to spike amount of post reconstitution or PR)
    rec.DryReconstitute.mean = PC.mean / PR.mean * 100, 
    rec.DryReconstitute.sd = sqrt((PC.sd / PC.mean)^2 + (PR.sd / PR.mean)^2) * rec.DryReconstitute.mean,
    
    # # add silica gel effect (relative to without, using spike level C)
    # sorbent.mean = C.mean  / `no silica gel.mean`, # (improvenment fold), with improvement > 1
    # sorbent.sd = sqrt((`no silica gel.sd` / `no silica gel.mean`)^2 + (C.sd / C.mean)^2) * sorbent.mean,
    
    # Overall recovery 
    rec.Overall.mean = B.mean / PR.mean * 100,
    rec.Overall.sd = sqrt((B.sd / B.mean)^2 + (PR.sd/ PR.mean)^2) * rec.Overall.mean,
    
    # matrix effect to ESI
    matrixESI.mean = PR.mean / `PR neat.mean` * 100,
    matrixESI.sd = sqrt((PR.sd / PR.mean)^2 + (`PR neat.sd` / `PR neat.mean`)^2) * matrixESI.mean,
    
    # Overall processing efficiency
    process.eff.mean = B.mean / `PR neat.mean` * 100,
    process.eff.sd = sqrt((B.sd / B.mean)^2 + (`PR neat.sd` / `PR neat.mean`)^2) * process.eff.mean) %>%
  
  select(compound, contains("rec"), contains("ESI"), contains("process"))

d.recovery.corrected
# colnames(d.recovery.corrected) = c("compound", str_c("crrct.", colnames(d.recovery.corrected)[-1]))
# d.recovery.corrected


# -------<>-------<>-------<>-------<>-------<>-------<>-------<>-------<>-------<>-------<>-------

x = d.recovery.corrected %>%
  select(compound, contains("mean")) %>%
  gather(-compound, key = step, value = mean) %>%
  mutate(step = str_remove(step, pattern = ".mean"))

y = d.recovery.corrected %>%
  select(compound, contains("sd")) %>%
  gather(-compound, key = step, value = sd) %>%
  mutate(step = str_remove(step, pattern = ".sd"))

d.recovery.corrected.tidy = x %>% left_join(y, by = c("compound", "step"))


# combine dataset of recovery corrected and not corrected
d.recovery.all = rbind(d.recovery.corrected.tidy %>% mutate(corrected = "ISTD-corrected") %>% ungroup(), 
                       d.recovery.tidy %>% mutate(corrected = "Absolute") %>% ungroup())

d.recovery.all$step = d.recovery.all$step %>% str_replace(pattern = "rec.extraction", replacement = "Extraction")
d.recovery.all$step = d.recovery.all$step %>% str_replace(pattern = "rec.EMR", replacement = "Enhanced matrix removal")
d.recovery.all$step = d.recovery.all$step %>% str_replace(pattern = "rec.DryReconstitute", replacement = "Dry & reconstitute")
d.recovery.all$step = d.recovery.all$step %>% str_replace(pattern = "rec.Overall", replacement = "Overall recovery")

d.recovery.all$step = d.recovery.all$step %>% str_replace(pattern = "matrixESI", replacement = "Matrix effect (ESI)")
d.recovery.all$step = d.recovery.all$step %>% str_replace(pattern = "process.eff", replacement = "Overall processing efficiency")


# ordered step
d.recovery.all$step = factor(d.recovery.all$step, levels = d.recovery.all$step %>% unique(), ordered = T)



# plot
dg = .4

plt.recovery.all = d.recovery.all %>% 
  ggplot(aes(x = corrected, y = mean)) +
  facet_wrap(~step, nrow = 2) +
  
  geom_boxplot(outlier.alpha = 0, fill = "grey") +
  
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd, color = compound),
                position = position_dodge(dg), 
                size = .2, width = 1) +
  
  
  geom_point(aes(color = compound),
             position = position_dodge(dg),
             size = 4.5, fill = "white", shape = 21) +
  
  geom_point(aes(color = compound, fill = compound),
             position = position_dodge(dg),
             size = 4.5, shape = 21, alpha = .2) +
  
  # geom_point(position = position_dodge(dg),
  #            shape = 21, size = 3, stroke = 1, alpha = .2) +
  
  coord_cartesian(ylim = c(0, 180)) +
  scale_y_continuous(breaks = seq(0, 200, by = 20)) +
  scale_color_manual(values = color.cmpd) +
  scale_fill_manual(values = color.cmpd) +
  theme(legend.position = "None", panel.grid = element_blank()) +
  labs(x = "  ") +
  
  # geom_text_repel(data = d.recovery.all$step %>% filter(mean > 110 | mean <= 60), 
  #                 aes(label = compound), position = position_dodge(dg), size = 3) +
  
  annotate("segment", x = .5, xend = 2.5, y = 120, yend = 120, 
           linetype = "dotted", size = .6, color = "firebrick") +
  annotate("segment", x = .5, xend = 2.5, y = 80, yend = 80, 
           linetype = "dotted", size = .6, color = "firebrick") +
  annotate("segment", x = .5, xend = 2.5, y = 100, yend = 100, 
           linetype = "dotted", size = .6, color = "darkgreen") +
  
  geom_text_repel(data = d.recovery.all %>% filter(mean < 50 | mean > 118 ),
                  aes(label = compound, color = compound), size = 5)
```


```{r, warning=F, message=F, fig.width=10, fig.height=10}
plt.recovery.all
```


```{r, warning=F, message=F}
# grid.arrange(plt.accuracy, plt.recovery.all, nrow = 2)
```


# Accumulated recovery
```{r, warning=F, message=F}

# Compound-stepwise recovery (accumulative effect) Not corrected
d.recovery.cummulated = d.recovery.all %>% filter(corrected == "Absolute" ) %>%
  # filter(compound == "4-HBA") %>%
  select(-sd, -corrected) %>%
  spread(key = step, value = mean)

d.recovery.cummulated = d.recovery.cummulated %>%
  mutate(Start = 100 + rnorm(n = nrow(d.recovery.cummulated), mean = 0, sd = 0), 
         # some random noise over 100 to avoid overlap
         
         Extraction.0 = Start,
         Extraction.1 = Extraction,
         
         EMR.0 = Extraction.1,
         EMR.1 = Extraction.1 * `Enhanced matrix removal` / 100,
         
         Dry.0 = EMR.1, 
         Dry.1 = EMR.1 * `Dry & reconstitute` / 100,
         
         matrixEffect.0 = Dry.1,
         matrixEffect.1 = Dry.1 * `Matrix effect (ESI)` / 100, 
         Detected = matrixEffect.1) %>%
  select(compound, Start, contains(".0"), contains(".1"), Detected) %>%
  gather(-compound, key = step, value = recovery.cummulated)


# make a copy for later marking the plot the position of accumulated recovery on the y axis (horizontal line)
ymark = d.recovery.cummulated[d.recovery.cummulated$step %>% str_detect(pattern = ".1"), ]
ymark$step = ymark$step %>% str_remove(pattern = ".1")
ymark = ymark %>% mutate(status = "Absolute")


d.recovery.cummulated$step = d.recovery.cummulated$step %>% str_remove(pattern = ".0")  
d.recovery.cummulated$step = d.recovery.cummulated$step %>% str_remove(pattern = ".1")  

d.recovery.cummulated$step = d.recovery.cummulated$step %>% 
  factor(levels = d.recovery.cummulated$step %>% unique(), ordered = T)


# add IS. label
d.recovery.cummulated = 
  d.recovery.cummulated %>% left_join(d.IS, by = "compound")

# d.recovery.cummulated %>% as.data.frame()

#<>-<>-<>-<>-<>-<>-<>-<>-<>-<>-<>-<>-<>-<>-<>-<>-<>-<>-

# Compound-stepwise recovery (accumulative effect) Corrected

# d.recovery.all %>% arrange(compound, step) %>% as.data.frame()


d.recovery.cummulated.corrected = d.recovery.all %>% 
  filter(corrected == "ISTD-corrected" ) %>%
  select(-sd, -corrected) %>%
  spread(key = step, value = mean)

d.recovery.cummulated.corrected = d.recovery.cummulated.corrected %>%
  mutate(Start = 100, 
         # may also consider adding some random noise over 100 to avoid overlap
         
         Extraction.0 = Start,
         Extraction.1 = Extraction,
         
         EMR.0 = Extraction.1,
         EMR.1 = Extraction.1 * `Enhanced matrix removal` / 100,
         
         Dry.0 = EMR.1, 
         Dry.1 = EMR.1 * `Dry & reconstitute` / 100,
         
         matrixEffect.0 = Dry.1,
         matrixEffect.1 = Dry.1 * `Matrix effect (ESI)` / 100, 
         Detected = matrixEffect.1) %>%
  select(compound, Start, contains(".0"), contains(".1"), Detected) %>%
  gather(-compound, key = step, value = recovery.cummulated)

# d.recovery.cummulated.corrected %>% as.data.frame()



# make a copy for later marking the plot the position of accumulated recovery on the y axis
ymark.corrected = d.recovery.cummulated.corrected[d.recovery.cummulated.corrected$step %>% str_detect(pattern = ".1"), ]
ymark.corrected$step = ymark.corrected$step %>% str_remove(pattern = ".1")
ymark.corrected = ymark.corrected %>% mutate(status = "ISTD-corrected")


d.recovery.cummulated.corrected$step = d.recovery.cummulated.corrected$step %>% str_remove(pattern = ".0")  
d.recovery.cummulated.corrected$step = d.recovery.cummulated.corrected$step %>% str_remove(pattern = ".1")  

d.recovery.cummulated.corrected$step = d.recovery.cummulated.corrected$step %>% 
  factor(levels = d.recovery.cummulated.corrected$step %>% unique(), ordered = T)

# d.recovery.cummulated.corrected %>% 
#   arrange(compound, step) %>% as.data.frame() # so far correct


# add IS. label
d.recovery.cummulated.corrected = 
  d.recovery.cummulated.corrected %>% left_join(d.IS, by = "compound")


# d.recovery.cummulated.corrected %>%
#   arrange(compound) %>% as.data.frame()



# Combine both dataset
d.recovery.cummulated.all = d.recovery.cummulated.corrected %>% mutate(status = "ISTD-corrected") %>%
  rbind(d.recovery.cummulated %>% mutate(status = "Absolute"))

# d.recovery.cummulated.all %>%
#   arrange(compound, status) %>% as.data.frame()


#plot
dg = .2

# plot 1
```


```{r, warning=F, message=F, fig.width=10}
d.recovery.cummulated.all %>%
  ggplot(aes(x = step, y = recovery.cummulated, color = compound)) +
  geom_line(aes(group = compound), position = position_dodge(dg), alpha = .7) +
  theme(legend.position = "None", 
        panel.grid = element_blank()) +
  scale_y_continuous(breaks = seq(0, d.recovery.cummulated.corrected$recovery.cummulated %>% max(), by = 10)) +
  facet_wrap(~status, nrow = 1) +
  scale_color_manual(values = color.cmpd) +
  labs(x = "  ", y = "Recovery (%)")
```


```{r, warning=F, message=F}
# plot 2
d.recovery.all$step = d.recovery.all$step %>% str_replace(pattern = "Enhanced matrix removal", replacement = "EMR")
d.recovery.all$step = d.recovery.all$step %>% str_replace(pattern = "Dry & reconstitute", replacement = "Dry")
d.recovery.all$step[d.recovery.all$step == "Matrix effect (ESI)"] = "matrixEffect"
# not knowing why but this following command not work....
d.recovery.all$step = d.recovery.all$step %>% str_replace(pattern = "Matrix effect (ESI)", replacement = "matrixEffect")


# combine accumulated recovery with stepwise recovery 
d.recovery.cummulated.all.withStepWise = 
  d.recovery.cummulated.all %>%
  left_join(d.recovery.all %>% rename(status = corrected), 
            by = c("compound", "step", "status")) %>%
  mutate(stepLossPercent = mean - 100)


# d.recovery.cummulated.all.withStepWise %>% select(compound, step, recovery.cummulated, status) %>% 
#   arrange(compound, status) %>%
#   as.data.frame()


# mark y position in the plot for accumulated recovery -- horizontal line
d.recovery.cummulated.all.withStepWise = 
  d.recovery.cummulated.all.withStepWise %>%
  left_join(rbind(ymark, ymark.corrected) %>% rename(ymark = recovery.cummulated), 
            by = c("compound", "step", "status"))


# mark y position in the plot for stepwise loss -- vertical line
d.recovery.cummulated.all.withStepWise = 
  d.recovery.cummulated.all.withStepWise %>%
  group_by(compound, status, step)  %>%
  summarise(lossMarkPosition = mean(recovery.cummulated)) %>%
  
  left_join(d.recovery.cummulated.all.withStepWise, by = c("compound", "status", "step"))



# now the main body has been built up, let's add the error band

# pp1 = d.recovery.all %>%
#   filter(corrected == "ISTD-corrected") %>% select(-corrected)
# 
# pp1.mean = pp1 %>% select(-sd) %>% spread(step, mean) %>%
#   select(compound, Extraction, EMR, Dry)
# names(pp1.mean) = c("compound", names(pp1.mean)[-1] %>% str_c(".mean"))
# 
# pp1.sd = pp1 %>% select(-mean) %>% spread(step, sd) %>%
#   select(compound, Extraction, EMR, Dry)
# names(pp1.sd) = c("compound", names(pp1.sd)[-1] %>% str_c(".sd"))
# 
# pp1.mean %>% left_join(pp1.sd, by = "compound") %>%
#   mutate(Extraction = Extraction.mean,
#          Extraction.errorBand = Extraction.sd,
#          
#          EMR = Extraction.mean * EMR.mean / 100,
#          EMR.errorBand = 
#          )


# re-compute the accumulated recovery (as cross validation) more importantly the errorband!!
myFunc1 = function(dataset){
  kkk1 = dataset %>% 
    mutate(Extraction = B.mean / PX.mean * 100,
           Extraction.errorBand = sqrt((B.sd / B.mean)^2 + (PX.sd / PX.mean)^2) * Extraction,
           
           EMR = B.mean / PC.mean * 100,
           EMR.errorBand = sqrt((B.sd / B.mean)^2 + (PC.sd / PC.mean)^2) * EMR,
           
           Dry = B.mean / PR.mean * 100, 
           Dry.errorBand = sqrt((B.sd / B.mean)^2 + (PR.sd / PR.mean)^2) * Dry,
           
           matrixEffect = B.mean / `PR neat.mean` * 100,
           matrixEffect.errorBand = sqrt((B.sd / B.mean)^2 + (`PR neat.sd` / `PR neat.mean`)^2) * matrixEffect) %>%
    
    select(-contains(".mean"), -contains(".sd")) %>% 
    ungroup()
  
  uu1 = kkk1 %>% select(compound, Extraction, EMR, Dry, matrixEffect) %>%
    gather(-compound, key = step, value = recovery.cummulated)
  
  uu2 = kkk1 %>% select(compound, contains("errorBand")) %>%
    gather(-compound, key = step, value = recovery.cummulated) 
  uu2$step = uu2$step %>% str_extract(one_or_more(WRD))
  uu2 = uu2 %>% rename(recovery.cummulated.errorBand = recovery.cummulated)
  
  kkk1 = uu1 %>% left_join(uu2, by = c("compound", "step"))
  return(kkk1)
}


qqq1 = myFunc1(dataset = d.area.steps %>% select(-C.mean, -C.sd, -contains("silica") ) ) %>%
  mutate(status = "Absolute")

qqq2 = myFunc1(dataset = d.area.corrected) %>%
  mutate(status = "ISTD-corrected")

d.recovery.cummulated.errorBand = qqq1 %>% rbind(qqq2, by = c("compound", "step")) %>%
  mutate(recovery.cummulated = recovery.cummulated %>% as.numeric(),
         recovery.cummulated.errorBand = recovery.cummulated.errorBand %>% as.numeric())


kkk1 = d.area.steps %>% select(-C.mean, -C.sd, -contains("silica")) %>%
  mutate(Extraction = B.mean / PX.mean * 100,
         Extraction.errorBand = sqrt((B.sd / B.mean)^2 + (PX.sd / PX.mean)^2) * Extraction,
         
         EMR = B.mean / PC.mean * 100,
         EMR.errorBand = sqrt((B.sd / B.mean)^2 + (PC.sd / PC.mean)^2) * EMR,
         
         Dry = B.mean / PR.mean * 100,
         Dry.errorBand = sqrt((B.sd / B.mean)^2 + (PR.sd / PR.mean)^2) * Dry,
         
         matrixEffect = B.mean / `PR neat.mean` * 100,
         matrixEffect.errorBand = sqrt((B.sd / B.mean)^2 + (`PR neat.sd` / `PR neat.mean`)^2) * matrixEffect) %>%
  
  select(-contains(".mean"), -contains(".sd")) %>%
  ungroup()

uu1 = kkk1 %>% select(compound, Extraction, EMR, Dry, matrixEffect) %>%
  gather(-compound, key = step, value = recovery.cummulated)

uu2 = kkk1 %>% select(compound, contains("errorBand")) %>%
  gather(-compound, key = step, value = recovery.cummulated)
uu2$step = uu2$step %>% str_extract(one_or_more(WRD))
uu2 = uu2 %>% rename(recovery.cummulated.errorBand = recovery.cummulated)

kkk1 = uu1 %>% left_join(uu2, by = c("compound", "step"))




# Combine with errorband dataset
d.recovery.cummulated.all.withStepWise = 
  d.recovery.cummulated.all.withStepWise %>% 
  left_join(d.recovery.cummulated.errorBand %>% rename(recovery.cummulated.homogenous = recovery.cummulated), 
            by = c("compound", "status", "step") )


# ordered factor
d.recovery.cummulated.all.withStepWise$step  = 
  factor(d.recovery.cummulated.all.withStepWise$step, 
         levels = c("Start", "Extraction", "EMR", "Dry", "matrixEffect", "Detected"), ordered = T)


# add numeric x axis to easy label the accumulated recovery
myX = seq(n_distinct(d.recovery.cummulated.all.withStepWise$step))  
names(myX) = d.recovery.cummulated.all.withStepWise$step %>% unique() %>% sort()
d.recovery.cummulated.all.withStepWise$numericX = myX[d.recovery.cummulated.all.withStepWise$step]



# Note that for "recovery.cummulated", each compound at each step has two values to form the vertical line in the plot, connecting the end of the prior step and the beginning of the following step. This term is only for purpse of drawing the line plot. For "recovery.cummulated.homogenous", its homogenous, it's the pure accumulated recovery at the end of the given step. It's used to draw the error band. The homogenous term is equal to one of the two recovery.cummulated values for a given compound at a given step (and also at the given status - IS corrected or not)


# plot
dg = 0
myalpha = .06

# define plot function

func.plotStepwiseRecovery = function(whichCompound) {
  
  dd.i = d.recovery.cummulated.all.withStepWise %>% filter(compound == whichCompound)
  
  dd.i %>% 
    ggplot(aes(x = numericX, y = recovery.cummulated, color = status)) +
    
    geom_line(aes(group = status), position = position_dodge(dg), size = 1.2) +
    
    theme(legend.position = c(.25, .2),
          legend.title = element_blank(),
          legend.text = element_text(size = 14),
          panel.grid = element_blank(),
          strip.text = element_text(size = 14),
          axis.text = element_text(colour = "black"),
          axis.title = element_text(colour = "black")
    ) +
    # scale_y_continuous(breaks = seq(0, max(dd.i$recovery.cummulated) + 10, by = 10)) +
    scale_x_continuous(breaks = myX) +
    
    facet_wrap(~compound, nrow = 3, scales = "free_y") +
    scale_color_brewer(palette = "Set1") +
    scale_fill_brewer(palette = "Set1") +
    labs(x = " ", y = "Recovery (%)") +
    
    
    # error band
    
    # extraction error band for accumulated recovery until end of extraction
    geom_ribbon(data = dd.i %>% filter(step == "Extraction"),
                aes(ymin = recovery.cummulated.homogenous - recovery.cummulated.errorBand,
                    ymax = recovery.cummulated.homogenous + recovery.cummulated.errorBand,
                    x = c(2,3, 2,3), fill = status), alpha = myalpha, position = position_dodge(dg), 
                linetype = "dotted", show.legend = F) +
    
    # EMR error band for accumulated recovery until end of EMR 
    geom_ribbon(data = dd.i %>% filter(step == "EMR"),
                aes(ymin = recovery.cummulated.homogenous - recovery.cummulated.errorBand,
                    ymax = recovery.cummulated.homogenous + recovery.cummulated.errorBand,
                    x = c(3,4, 3,4), fill = status), alpha = myalpha, position = position_dodge(dg),  
                linetype = "dotted", show.legend = F) +
    
    # Dry and recovery error band for accumulated recovery until end of reconstitution
    geom_ribbon(data = dd.i %>% filter(step == "Dry"),
                aes(ymin = recovery.cummulated.homogenous - recovery.cummulated.errorBand,
                    ymax = recovery.cummulated.homogenous + recovery.cummulated.errorBand,
                    x = c(4,5, 4,5), fill = status), alpha = myalpha, position = position_dodge(dg), 
                linetype = "dotted", show.legend = F) +
    
    # Matrix effect error band for accumulated recovery until end of ESI
    geom_ribbon(data = dd.i %>% filter(step == "matrixEffect"),
                aes(ymin = recovery.cummulated.homogenous - recovery.cummulated.errorBand,
                    ymax = recovery.cummulated.homogenous + recovery.cummulated.errorBand, 
                    x = c(5,6, 5,6), fill = status), alpha = myalpha, position = position_dodge(dg), 
                linetype = "dotted", show.legend = F) +
    
    # Add annotation
    
    # stepwise loss
    geom_label(aes(label = round(stepLossPercent, 1), y = lossMarkPosition), 
               size = 5.5, label.padding = unit(0.12, "lines"),
               # fontface = "bold",
               position = position_dodge(dg)) +
    
    # accumulated recovery
    geom_label(aes(label = round(ymark, 1), 
                   fill = status, 
                   y = ymark, x = numericX + .5), 
               fontface = "bold", 
               size = 5.5,label.padding = unit(0.12, "lines"),
               color = "white")  
  
}

y1 = func.plotStepwiseRecovery(whichCompound =  "RK")
y1
y2 = func.plotStepwiseRecovery(whichCompound =  "ROH")
y3 = func.plotStepwiseRecovery(whichCompound =  "RK-Me")
y4 = func.plotStepwiseRecovery(whichCompound =  "3-HBA")

plt.streamLine = plot_grid(y1, y2,  nrow = 2)


# y3 = plot_grid(yy, plt.accuracy, nrow = 1, rel_widths = c(7, 4))
```

# Accuracy variance vs.background
```{r, warning=F, message=F}
# accuracy variance with background interference
plt.accuracyVariance.vs.background = d.ABCD %>% select(compound, contains("Accuracy"), Name, conc.expected) %>% 
  left_join(d.conc.blk, by = "compound") %>%
  mutate(ratio.spk.vs.blk = conc.expected / conc.blk.mean) %>%
  
  
  ggplot(aes(x = ratio.spk.vs.blk, y = Accuracy.sd, color = Name)) +
  geom_point(size = 2, alpha = .5) +
  geom_point(size = 2, shape = 21, fill = NA) +
  geom_smooth(method = "lm",  show.legend = F, aes(fill = Name), alpha = .1) +
  scale_y_log10() +
  scale_x_log10() +
  
  annotation_logticks(sides = "lb") +
  scale_color_brewer(palette = "Set1") + 
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = "right",
        panel.grid = element_blank(),
        legend.text = element_text(size = 13))

plt.accuracyVariance.vs.background
```


```{r, warning=F, message=F}
plt.accuracyAnalysis = plot_grid(
  plt.accuracy, ggplot() + theme_void(), plt.accuracyVariance.vs.background, 
  nrow = 1, rel_widths = c(5, 1.5, 4))




plt.Validation.Accuracy.Recovery = plot_grid(
  
  plot_grid(plt.recovery.all, plt.streamLine, rel_widths = c(6, 3)), 
  plt.accuracyAnalysis,
  # nrow = 3,
  # rel_heights = c(3, 3.6, 3)
  
  nrow = 2,
  rel_heights = c(2.2, 1)
)
```

# Validation all plots
```{r, warning=F, message=F, fig.width=16, fig.height=18}
plt.Validation.Accuracy.Recovery
# 15.3 X 18
```


```{r, warning=F, message=F}
```

# Sorbent revisit2
#### The effect of biomatrices and sorbent on analytes' recovery
```{r, warning=F, message=F}
# ----Compare with the prior without sorbent in presence of biomatrices
# the first formal validation study that renders low recovery
mypath = "/Users/Boyuan/Desktop/My publication/17. EMR Adipose RK metabolites/Archived/EMR_all data.xlsx"


# peak area 
d.area = read_excel(mypath, sheet = "peak area")

d.area.copy = d.area

d.area = d.area %>% filter(Name %in% c("PRneat", "B", "PX", "PC", "PR", "BLK" )) %>%
  gather(-c(`Data File`, Name, `Acq. Date-Time`), key = compound, value = area)

d.area.blk = d.area %>%
  filter(Name == "BLK") %>% group_by(compound) %>%
  summarise(area.blk.mean = mean(area), area.blk.std = sd(area))

d.area.spk = d.area %>% filter(Name != "BLK") %>% group_by(compound, Name) %>% 
  summarise(area.mean = mean(area),
            area.sd = sd(area)) %>%
  # subtract blank
  left_join(d.area.blk, by = "compound") %>%
  mutate(area.spk = area.mean - area.blk.mean,
         area.spk.sd = sqrt( area.sd ^2 + area.blk.std ^ 2 )) %>%
  select(compound, Name, contains("spk"))

# tidy up
x = d.area.spk %>% select(-area.spk.sd) %>%
  spread(key = Name, value = area.spk)

y = d.area.spk %>% select(-area.spk) %>%
  spread(key = Name, value = area.spk.sd)
colnames(y) = c("compound", colnames(y)[-1] %>% str_c(".sd"))
y

d.area.spk.net = left_join(x, y, by = "compound")

# compute analyte loss at each step
d.area.spk.net = d.area.spk.net %>%
  mutate(rec.extract = B / PX * 100, 
         rec.cleanup = PX / PC * 100, 
         
         rec.dryConcentrate = PC / PR * 100,
         rec.dryConcentrate.sd = ((PC.sd / PC)^2 + (PR.sd / PR)^2) * rec.dryConcentrate )


d.area.spk.net.copy = d.area.spk.net %>% 
  select(compound, contains("rec")) %>%
  select(-contains("sd")) %>%
  gather(-1, key = step, value = recovery) 



# Combine the recovery with and without sorbent in presence of biomatrices
# data from the first validation study (the one that failed due to low recovery at drying step without sorbent)
k1 = d.area.spk.net %>% 
  select(compound, contains("rec.dryConcentrate")) %>%
  mutate(sorbent = "with biomatrix, no sorbent") %>% ungroup() 

# data from the second validation study conducted with sorbent of C18 ca. 2~5 mg/600 uL ACN extract
k2 = d.recovery.all %>%
  filter(step == "Dry") %>% filter(corrected == "Absolute") %>%
  select(compound, mean, sd) %>%
  mutate(sorbent = "with biomatrix, with C18") %>%
  rename(rec.dryConcentrate = mean, 
         rec.dryConcentrate.sd = sd) %>% ungroup()

d.recoveryCompare.sorbentWithBiomatrices = k1 %>% rbind(k2)


# combine with when having biomatrix, with or without sorbent
d.recoveryCompare.sorbentWithBiomatrices = 
  
  # no biomatrix, no sorbent
  d.opt.sorbent.recovery %>%
  filter(sorbent == "Sorbent free") %>% ungroup() %>%
  select(compound, recovery.mean, recovery.sd) %>%
  rename(rec.dryConcentrate = recovery.mean,
         rec.dryConcentrate.sd  = recovery.sd) %>%
  mutate(sorbent = "no biomatrix, no sorbent") %>%
  
  rbind(d.recoveryCompare.sorbentWithBiomatrices)


# ordered matrix-sorbent condition
d.recoveryCompare.sorbentWithBiomatrices$sorbent = 
  d.recoveryCompare.sorbentWithBiomatrices$sorbent %>% 
  factor(levels = d.recoveryCompare.sorbentWithBiomatrices$sorbent %>% unique(), ordered = T)



#
# plot
# arrange compound from low to high recovery
jjj = d.recoveryCompare.sorbentWithBiomatrices %>% 
  group_by(compound) %>%
  summarise(mean.rec = mean(rec.dryConcentrate)) %>%
  arrange(mean.rec)
cmpd.ordered.dryRecovery = jjj$compound %>% unique()

d.recoveryCompare.sorbentWithBiomatrices$compound = 
  d.recoveryCompare.sorbentWithBiomatrices$compound %>%
  factor(levels = cmpd.ordered.dryRecovery, ordered = T)


# now plot!

dg = .6
plt.recoveryCompare.sorbentWithBiomatrices =
  
  d.recoveryCompare.sorbentWithBiomatrices %>%
  
  filter(rec.dryConcentrate < 120) %>% # 4-HPPA, no biomatrix no sorbent the recoveryca. 150 outlier remove it
  
  ggplot(aes(x = compound, y = rec.dryConcentrate, fill = sorbent)) +
  geom_bar(stat = "identity", position = position_dodge(dg), alpha = 1, width = .7) +
  theme(panel.grid = element_blank(),
        legend.position = "bottom",
        legend.text = element_text(size = 11),
        legend.title = element_blank(),
        axis.text = element_text(size = 11)) +
  scale_y_continuous(breaks = seq(0, 120, 20)) +
  
  geom_errorbar(aes( ymin = rec.dryConcentrate - rec.dryConcentrate.sd,
                     ymax = rec.dryConcentrate + rec.dryConcentrate.sd,
                     color = sorbent),
                position = position_dodge(dg), width = .5) +
  coord_flip() +
  labs(y = "Recovery (%) ") +
  scale_color_manual(values = c("snow4", "tomato", "skyblue")) +
  scale_fill_manual(values = c("snow4", "tomato", "skyblue"))
```


```{r, warning=F, message=F, fig.height=8}
# plt.recoveryCompare.sorbentWithBiomatrices
```


```{r, warning=F, message=F, fig.width=12, fig.height=10}
# plot sorbent rsult
plot_grid(plt.sorbentOptimization,
          plt.recoveryCompare.sorbentWithBiomatrices, 
          nrow = 1, rel_widths = c(1.5, 1))
#  big screen 14.7 X 8.16
```

# Sorbent revist-3
#### Validation at level C,  50 ng/mL injection conc.
```{r, warning=F, message=F}
# Check the effect of with vs. without sorbent validated at level C， i.e. about 50 ng/mL
d.area = read_excel(path, sheet = "peak area")
gg = d.area %>%
  filter(Name %in% c("C", "no silica gel")) %>%
  select(-c(`Acq. Date-Time`, `Data File`)) %>%
  gather(-Name, key = compound, value = resp) %>%
  group_by(Name, compound) %>%
  summarise(resp.mean = mean(resp),
            resp.sd = sd(resp))


gg = gg %>%
  group_by(compound) %>%
  mutate(resp.mean.normalize = resp.mean / mean(resp.mean),
         resp.sd.normalize = resp.sd / mean(resp.mean)) %>%
  select(Name, compound, contains("normalize")) 

mj = gg %>% filter(Name == "C") %>%
  arrange(resp.mean.normalize)


gg$compound = factor(gg$compound, levels = rev(mj$compound), ordered = T)
```


```{r, warning=F, message=F, fig.width=8}
gg %>% filter(compound != "4-HPPA") %>%
  ggplot(aes(x = compound, y = resp.mean.normalize, color = Name, fill = Name)) +
  
  geom_bar(stat = "identity", alpha = .6, position = position_dodge(.6)) +
  
  geom_errorbar(aes(ymin = resp.mean.normalize - resp.sd.normalize,
                    ymax = resp.mean.normalize + resp.sd.normalize), 
                width = .3, position = position_dodge(.6)) +
  # coord_flip() +
  labs(y = "Normalized peak area") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid = element_blank())
```

# Check numbers for paper writing
```{r, warning=F, message=F}
# -- check some numbers for paper writing

d.checkProcessingEfficency.vs.accuracy.B.Level = d.ABCD %>% filter(Name == "B") %>% 
  select(compound, Accuracy.mean) %>% 
  
  left_join(d.recovery.all %>% filter(step == "Overall processing efficiency") %>%
              filter(corrected == "ISTD-corrected"), by = "compound") %>%
  select(compound, Accuracy.mean, mean) %>%
  rename(ProcessingEfficiency.corrected = mean) %>%
  mutate(ratio = Accuracy.mean / ProcessingEfficiency.corrected) %>%
  arrange(ratio)

d.checkProcessingEfficency.vs.accuracy.B.Level$compound = 
  factor(d.checkProcessingEfficency.vs.accuracy.B.Level$compound, 
         levels = d.checkProcessingEfficency.vs.accuracy.B.Level$compound, 
         ordered = T)

d.checkProcessingEfficency.vs.accuracy.B.Level %>% 
  ggplot(aes(x = compound, y = ratio)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_segment(aes(x = 1, xend = 26, y = 1, yend = 1))



d.recovery.all$step %>% unique()
px = d.recovery.all %>% 
  filter(step == "Overall processing efficiency") %>%
  filter(corrected == "ISTD-corrected") %>%
  arrange(mean) %>%
  as.data.frame() 
# px
px$mean %>% hist(breaks = 26)



kkkkk = d.recovery.all %>% 
  filter(step == "Overall processing efficiency") %>%
  select(-step, -sd) %>%
  spread(key = corrected, value = mean) %>%
  mutate(correctAmoung = `ISTD-corrected` - Absolute)
kkkkk$correctAmoung %>% hist(breaks = 26)   






pp = d.recoveryCompare.sorbentWithBiomatrices %>%
  select(-contains(".sd")) %>%
  spread(sorbent, rec.dryConcentrate) %>%
  mutate(sorb.improv = `with biomatrix, with C18` - `with biomatrix, no sorbent`) %>%
  arrange(sorb.improv) %>%
  as.data.frame()
# pp

```

